---
title: "Homework 02"
---


## TidyTuesday Section (optional)

::: {.callout-important title="Instructions"}

You can count work on this week's TidyTuesday toward the exceptional work required for an A in the Homework component.

:::


Explore the week's [TidyTuesday](https://github.com/rfordatascience/tidytuesday) challenge. Develop a research question, then answer it through a short data story with [effective visualization(s)]({{< var effective-viz-url >}}).  Provide sufficient background for readers to grasp your narrative.

What is the difference of the proportion of protein for different Chinese chicken cuisine ? 

```{r}
library(tidyverse)
library(ggplot2)
library(stringr)
library(ggrepel)
```

```{r}
cuisine <- read.csv("./cuisines.csv")
```

```{r}
cuisine_new <- cuisine |>
  filter(country == c("Chinese", "Brazilian", "Indian", "American", "Turkish", "Japanese"))|>
  group_by(country) |>
  summarize(max_rating = max(avg_rating, na.rm = TRUE),
            min_rating = min(avg_rating, na.rm = TRUE),
            mean_rating = mean(avg_rating, na.rm = TRUE),
            .groups = "drop",
            top_dish = name[which.max(avg_rating)])
```

```{r}
cuisine_plot <- ggplot(cuisine_new) +
  geom_linerange(aes(x = country, ymin = min_rating, ymax = max_rating), linewidth = 20, color = "pink")+
  geom_text_repel(aes(x = country, y = max_rating, label = top_dish, force = 500, force_pull = 500))+
  geom_point(aes(x = country, y = max_rating, text = top_dish), shape = 25, fill = "red", size = 2)+
  labs(
    title = "General Ratings Among Five Countries Cuisines",
    x = "Country",
    y = "Rating"
  ) +
    theme_classic()
cuisine_plot
  
```

