---
title: "Homework 02"
---


## TidyTuesday Section (optional)

::: {.callout-important title="Instructions"}

You can count work on this week's TidyTuesday toward the exceptional work required for an A in the Homework component.

:::


Explore the week's [TidyTuesday](https://github.com/rfordatascience/tidytuesday) challenge. Develop a research question, then answer it through a short data story with [effective visualization(s)]({{< var effective-viz-url >}}).  Provide sufficient background for readers to grasp your narrative.

What is the difference of the proportion of protein for different Chinese chicken cuisine ? 

```{r}
library(tidyverse)
library(ggplot2)
library(stringr)
library(ggrepel)
```

```{r}
cuisine <- read.csv("./cuisines.csv")
```

```{r}
cuisine_new <- cuisine |>
  filter(country == c("Chinese", "Brazilian", "Indian", "American", "Turkish", "Japanese"))|>
  group_by(country) |>
  summarize(max_rating = max(avg_rating, na.rm = TRUE),
            min_rating = min(avg_rating, na.rm = TRUE),
            mean_rating = mean(avg_rating, na.rm = TRUE),
            .groups = "drop",
            top_dish = name[which.max(avg_rating)])
```

```{r}
cuisine_plot <- ggplot(cuisine_new) +
  geom_linerange(aes(x = country, ymin = min_rating, ymax = max_rating), linewidth = 20, color = "pink")+
  geom_text_repel(aes(x = country, y = max_rating, label = top_dish, force = 500, force_pull = 500))+
  geom_point(aes(x = country, y = max_rating, text = top_dish), shape = 25, fill = "red", size = 2)+
  labs(
    title = "General Ratings Among Five Countries Cuisines",
    x = "Country",
    y = "Rating"
  ) +
    theme_classic()
cuisine_plot
  
```

> My question is how does the general ratings different among these five countries cuisines, since they are all my friends home cuisines. These recipes all came from this one website called allrecipes, the rating is range from 0-5 stars. And for me, surprisingly, all five countries have similar top ratings, and the range is generally from 4 to 5 stars, except for this one outlier which is indian cuisine, which suprising have one star and five star review recipes. 
